AGU Timeseries

### 2017-12-05 r_scripts/analysis/agu_saba_timeseries.Rmd
Made figure for jeff and kate.

# time sereis of saba
saba_ts <- saba %>% 
  group_by(MSA, date) %>% 
  summarise(saba_n = n()) %>% 
  mutate(metroarea = case_when(
    MSA == 13460 ~ "Bend-Redmond",
    MSA == 18700 ~ "Corvallis",
    MSA == 21660 ~ "Eugene-Springfield",
    MSA == 32780 ~ "Medford",
    MSA == 38900 ~ "Portland-Vancouver-Beaverton",
    MSA == 41420 ~ "Salem",
    MSA == 41999 ~ "Not in MSA"),
    day = weekdays(date),
    weekend = ifelse(day %in% c("Saturday", "Sunday"), "Weekend", "Weekday"))

@@

```{r msa plot}
plot <- ggplot(saba_ts, 
    aes(x=date, y=saba_n, group = weekend, color = weekend)) +
  geom_point() +
  geom_hline(data=saba_center_stats, aes(yintercept=saba_mean, group=weekend,
    color=weekend)) +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  facet_wrap(~metroarea) +
  ylab("SABA Count") +
  xlab("Date") +
  theme_minimal()

@@

@@

```{r medford count timeseries}
# medford time series
ts_plot <- ggplot(medford_ts, 
  aes(x=date, y=saba_n, group=weekend, color=weekend)) +
  geom_point() +
  geom_hline(data=medford_center_stats, aes(yintercept=saba_mean, color=weekend)) +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylim(0, 50) +
  ylab("SABA Count") +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme_minimal()

@@

```{r gwr smk timeseries}
# jackson plot
jackson_pm <- county_pm_df %>% 
  filter(county == "Jackson") %>% 
  mutate(smk10 = ifelse(geo_wt_pm > 10, 1, 0),
    fire_period = ifelse(date >= "2013-07-26" & date <= "2013-09-03",1,0))

# plot 
pm_plot <- ggplot(jackson_pm, aes(x=date, y=geo_wt_pm)) +
  geom_point(color = "#000046") +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylab(expression(paste("PM"[2.5], "  ", mu, "g/", m^3, sep=""))) +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme_minimal()

@@

>>>>

# time sereis of saba
saba_ts <- saba %>% 
  group_by(MSA, date) %>% 
  summarise(saba_n = n()) %>% 
  mutate(metroarea = case_when(
    MSA == 13460 ~ "Bend-Redmond",
    MSA == 18700 ~ "Corvallis",
    MSA == 21660 ~ "Eugene-Springfield",
    MSA == 32780 ~ "Medford",
    MSA == 38900 ~ "Portland-Vancouver-Beaverton",
    MSA == 41420 ~ "Salem",
    MSA == 41999 ~ "Not in MSA"),
    day = weekdays(date),
    weekend = ifelse(day %in% c("Saturday", "Sunday"), "Weekend", "Weekday"),
    week_n = lubridate::floor_date(date, unit = "week"))

@@

### Figure 1: Time-series of short-acting beta-2 agonists over the metropolitan statistical areas for the state of Oregon.

```{r msa plot}
plot <- ggplot(saba_ts, 
    aes(x=date, y=saba_n, group = weekend, color = weekend)) +
  geom_point() +
  geom_hline(data=saba_center_stats, aes(yintercept=saba_mean, group=weekend,
    color=weekend)) +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  facet_wrap(~metroarea) +
  ylab("SABA Count") +
  xlab("Date") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_line(colour="grey"),
    #panel.grid.minor = element_blank(),
    legend.position = "bottom",
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"))

@@

### Figure 2: Time-series of SABA limited to Medford MSA.

@@

### Figure 3: Time-series of SABA Z-Score.

```{r medford count timeseries}
# medford time series
ts_plot <- ggplot(medford_ts, 
  aes(x=date, y=saba_n, group=weekend, color=weekend)) +
  geom_point() +
  geom_hline(data=medford_center_stats, aes(yintercept=saba_mean, color=weekend)) +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylim(0, 50) +
  ylab("SABA Count") +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom")

@@

### Figure 4: Truncated SABA count

I'm truncating to July to October.

```{r trunc saba ts}
jul_oct_medford <- medford_ts %>% 
  filter(date >= "2013-07-01" & date <= "2013-10-01")

jul_oct_central_vals <- medford_ts %>% 
  group_by(weekend) %>% 
  summarise(saba_med = median(saba_n))

# medford time series
saba_plot <- ggplot(jul_oct_medford, 
  aes(x=date, y=saba_n, group=weekend, color=weekend)) +
  geom_point() +
  #geom_hline(data=jul_oct_central_vals , aes(yintercept=saba_med, color=weekend)) +
  scale_color_manual(name = "Days of Week", values = c("#ec38bc", "#7303c0")) +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylim(0, 40) +
  ylab("SABA Count") +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom",
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"))

saba_plot
```

### Figure 5: Time-series of GWR PM~2.5~ values.

```{r gwr smk timeseries}
# jackson plot
jackson_pm <- county_pm_df %>% 
  filter(county == "Jackson") %>% 
  mutate(smk10 = ifelse(geo_wt_pm > 10, 1, 0),
    fire_period = ifelse(date >= "2013-07-26" & date <= "2013-09-03",1,0)) %>% 
  filter(date >= "2013-07-01" & date <= "2013-10-01")

# plot 
pm_plot <- ggplot(jackson_pm, aes(x=date, y=geo_wt_pm)) +
  geom_point(color = "#000046") +
  geom_vline(xintercept = as.numeric(as.Date(c("2013-07-26", "2013-09-03"), 
      "%Y-%m-%d")),
    linetype=3, color = "#191654") +
  ylab(expression(paste("PM"[2.5], "  ", mu, "g/", m^3, sep=""))) +
  xlab("Date") +
  #ggtitle("Medford SABA Counts by Date") +
  theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank())

@@

### Figure 6: Binding truncated SABA counts and PM~2.5~

```{r trunc plot}
cowplot::plot_grid(pm_plot, saba_plot,
  #labels = c("PM2.5 Exposure", "Rescue Inhaler"), 
  nrow = 2, align = "v")
```
